Examining Crowd Work and Gig Work Through The Historical Lens of Piecework

### 2017-05-03 chi/piecework/presentation/complexity.tex
complexity lit review done?

@@

\begin{frame}{Complexity}
  What kinds of problems do we mean when we talk about complexity?
  \begin{mystepwiseitemize}
    \item Can crowds improve existing works?~\cite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986}
    \item Can crowds critique designs?~\cite{yuanAlmost}
    \item Can crowds create things from whole cloth?~\cite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}
  \end{mystepwiseitemize}
\end{frame}

\begin{frame}{What does the crowdsourcing literature say?}
\begin{columns}
  \begin{column}{0.5\textwidth}
    \begin{itemize}
      \item Build complexity into the process
      \begin{mystepwiseitemize}
        \item Apply CS methods to people~(\textcite{crowdForgeKittur})
        \item see other stuff
      \end{mystepwiseitemize}
    \end{itemize}
  \end{column}

  \begin{column}{0.5\textwidth}
    \begin{figure}
    % \visible<2>{
    \includegraphics[width=0.7\textwidth]{figures/mapReduce.png}
    % }
    % \visible<1>{

    \includegraphics[width=0.7\textwidth]{figures/complexity/cw_literature/scribe.png}
    % }
    \end{figure}
  \end{column}

@@

\begin{frame}{What does the piecework literature say?}
    something even more insightful, I'm sure!
\end{frame}

@@

%   \end{column}
%   \end{columns}
\end{frame}
\end{document}

>>>>

\onlyinsubfile{
\bibliography{references}
  % \beamertemplategridbackground[1mm]
}

@@

\begin{frame}{Complexity}
  What kinds of problems do we mean when we talk about complexity?
  \begin{itemize}
    \item<1-> Can crowds help you write something?

    \scriptsize{
      \textcite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986,Nebeling:2016:WCW:2858036.2858169}
    }
    \item<2-> Can crowds critique novel designs?

    \scriptsize{
      \textcite{yuanAlmost,fuge2014analysis}
    }
    \item<3-> Can crowds create things from whole cloth?

    \scriptsize{
      \textcite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}
    }
  \end{itemize}
\end{frame}

\begin{frame}{What does the crowdsourcing literature say?}
\begin{columns}
  \begin{column}{0.6\textwidth}
    \begin{itemize}
      \item Build complexity into the process
      \begin{mystepwiseitemize}
        \item Apply CS methods to people

        \scriptsize{
          \textcite{crowdForgeKittur}
        }
        \item Humans as computational units

        \scriptsize{
          \textcite{Lasecki:2014:LSR:2661334.2661352}
        }
        \item Crowdsourcing workflows as function state machines

        \scriptsize{\textcite{latoza2014microtask}
        }
      \end{mystepwiseitemize}
    \end{itemize}
  \end{column}

  \begin{column}{0.4\textwidth}
    \begin{figure}
    \only<1>{
    \includegraphics[width=\textwidth]{figures/complexity/cw_literature/mapReduce.png}
    }

    \only<2>{
      \includegraphics[width=\textwidth]{figures/complexity/cw_literature/scribe.png}
    }
    \only<3>{
      \begin{adjustbox}{max totalsize={\textwidth}{\textheight},center}
      \begin{tikzpicture}[->,>=stealth',auto, node distance=4cm,
                          thick,transform shape]
        \node[state,label=above:{Described}]                       (A)                    {};
        \node[state,label=left:{Described written}]                (B) [below of=A]       {};
        \node[state,label=left:{Run Tests}]                        (D) [below of=B]       {};
        \node[state,label=below:{Described written buggy}]         (C) [below right of=D] {};
        \node[state,label=below:{Described written buggy}]         (E) [below left of=D]  {};

        \path[every node/.style={sloped,anchor=south}]
              (A) edge              node {Write description} (B)
              (B) edge [loop right] node {Edit code} (B)
                  edge              node {Edit code} (D)
              (D) edge              node {} (C)
                  edge              node {} (E)
              (C) edge [bend right] node {Debug} (B)
                  edge [bend right] node {Debug} (D)
              (E) edge [bend right] node {Edit Code} (D)
                  edge [bend left]  node {Edit Code} (B);
        \end{tikzpicture}
        \end{adjustbox}
    }
    \end{figure}
  \end{column}

@@

\begin{frame}{What does the piecework literature say?}
    % \begin{itemize}
      % \item
      George Airy (astronomer) used a very similar approach~\cite{grier2013computers}
      % \begin{itemize}
      %   \item Break work up so that individual tasks are simple

      %   Make the process \textit{itself} compile the final product.
      % \end{itemize}
    % \end{itemize}
    % \vspace{0mm}
    \begin{columns}
    \begin{column}{0.6\textwidth}
      \begin{figure}
        \includegraphics[width=\textwidth]{figures/photo/Greenwich-Observatory-computing-staff-1902.jpg}
      \end{figure}
    \end{column}

    \begin{column}{0.4\textwidth}
      \begin{itemize}
        \item Employed computers
        \item 13--20 years old
        \item Overworked
        \item Underpaid
        \item Could be fired at will
      \end{itemize}
    \end{column}
    \end{columns}

\end{frame}

\begin{frame}{George Airy --- whiz kid}
    Airy built complexity into the process, assigning \textit{human computers} 
    to compute, verify, and correct the right ascension and declination of stars.

    \begin{figure}
    \includegraphics[width=\textwidth]{figures/complexity/pw_literature/airy.png}
    \end{figure}
\end{frame}

\begin{frame}{Humble origins}
  First implementations of piecework:
  \visible<2->{farms}\visible<3->{, textiles}\visible<4->{, and other low--skill labor.}
  \begin{columns}

  \visible<2->{
  \begin{column}{0.3\textwidth}
  \begin{figure}
  \includegraphics[width=\textwidth]{figures/idk.png}
  \end{figure}
  \end{column}
  }

  \visible<3->{
  \begin{column}{0.3\textwidth}
  \begin{figure}
  \includegraphics[width=\textwidth]{figures/pieceworkers.jpg}
  \end{figure}
  \end{column}
  }

  \visible<4->{
  \begin{column}{0.3\textwidth}
  \begin{figure}
  \includegraphics[width=\textwidth]{figures/photo/Matchgirl_strikers.PNG}
  \end{figure}
  \end{column}
  }
  \end{columns}
\end{frame}

\begin{frame}{Ford}
    Fordism

\end{frame}

@@

%   \end{column}
%   \end{columns}
\end{frame}

\onlyinsubfile{
  \printbibliography{}
}
\end{document}

### 2017-05-03 chi/piecework/presentation/presentation.tex
complexity lit review done?

\usepackage{tikz,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles}
\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
\usepackage[export]{adjustbox}
\usepackage[en-US]{datetime2}
\bibliography{references}
\usetheme{metropolis}
\beamertemplatenavigationsymbolsempty
\renewcommand\textbullet{\ensuremath{\bullet}}
\definecolor{stanfordRed}{HTML}{8C1515}
% \mode<presentation>{
  % \usetheme{metropolis}
  \usecolortheme{seagull}

@@

\newcommand{\onlyinsubfile}[1]{#1}
\begin{document}
\renewcommand{\onlyinsubfile}[1]{}
\begin{frame}
\titlepage
\end{frame}
% \setbeamercovered{transparent}

\subfile{intro.tex}

\begin{frame}[standout]
    What is the future of work?
\end{frame}

\begin{frame}{What is the future of work?}

\end{frame}

\begin{frame}{Introduction}
  We hope to provide:
      \begin{itemize}
        \item A useful ontological lens for making sense of crowdsourcing and gig work (which we collectively call ``\textit{on--demand work}'') as a resurgence of \textit{piecework}.
        \item A method for making sense of contemporary phenomena through \textit{historical analysis}.
      \end{itemize}
\end{frame}

\begin{frame}{A case for comparative historical analysis}
\begin{itemize}
  \item Historical analysis isn't new
  \begin{itemize}
    \item In general (see~\textcite{rosenberg1994exploring,rosenberg1982inside})
    \item In HCI (see~\textcite{Wyche2006,bodker1993historical})
  \end{itemize}
  % \item Still can be useful.
\end{itemize}
\end{frame}

>>>>

\usepackage{tikz,pgf,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles,chronosys,cancel}
\usetikzlibrary{arrows,automata}
\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
\usepackage[export]{adjustbox}
\usepackage[en-US]{datetime2}
\bibliography{references}
\usetheme{metropolis}
\beamertemplatenavigationsymbolsempty
\renewcommand\textbullet{\ensuremath{\bullet}}
\definecolor{stanfordRed}{HTML}{8C1515}
% \mode<presentation>{
  % \usetheme{metropolis}
  \usecolortheme{seagull}

\usepackage{pgfcalendar}
\usepackage{datenumber,xparse,fp}
\usetikzlibrary{arrows.meta,backgrounds,fixedpointarithmetic}
\makeatletter
\ExplSyntaxOn
\tl_new:N \l_chronos_dateformat_tl
\tl_set:Nn \l_chronos_dateformat_tl { Y/m/d }
% YY yn lle YYYY
\cs_new_protected_nopar:Npn \chronos_year_shorten:n #1
{
  \chronos_year_shorten_aux:w #1 \q_stop % expl3 manuaal, 46
}
\cs_new_protected_nopar:Npn \chronos_year_shorten_aux:w #1 #2 #3 #4 \q_stop
{
  #3 #4
}
\cs_generate_variant:Nn \chronos_year_shorten:n { V , c }
% dangos dyddiadau
\cs_new_protected_nopar:Npn \chronos_show_date:n #1
{
  \tl_map_inline:Nn \l_chronos_dateformat_tl
  {
    \str_case:nnF { ##1 }
    {
      { a } { \pgfcalendarweekdayshortname{\thechronos@weekday} }
      { A } { \pgfcalendarweekdayname{\thechronos@weekday} }
      { b } { \pgfcalendarmonthshortname{\csname chronos@#1month\endcsname} }
      { B } { \pgfcalendarmonthname{\csname chronos@#1month\endcsname} }
      { d } { \csname chronos@#1day\endcsname }
      { m } { \csname chronos@#1month\endcsname }
      { y } { \chronos_year_shorten:c { chronos@#1year } }
      { Y } { \csname chronos@#1year\endcsname }
      { @ } { ~ }
    }
    {
      ##1
    }
  }
}
% user interface
\NewDocumentCommand \chronossetdateformat { m }
{
  \tl_set:Nn \l_chronos_dateformat_tl { #1 }
  \tl_replace_all:Nnn \l_chronos_dateformat_tl { ~ } { @ }
}
% for pgf/tikz convenience
\NewDocumentCommand \chronos@showdate { m }
{
  \pgfcalendarjuliantoweekday{\csname thechronos@#1date\endcsname}{\c@chronos@weekday}%
  \chronos_show_date:n { #1 }
}
\ExplSyntaxOff
\newcounter{chronos@date}
\newcounter{chronos@startdate}
\newcounter{chronos@enddate}
\newcounter{chronos@startyear}
\newcounter{chronos@endyear}
\newcounter{chronos@yeardate}
\newcounter{chronos@thingdate}
\newcounter{chronos@otherthingdate}
\newcounter{chronos@weekday}
\newif\ifchronos@marks
\chronos@marksfalse
\newif\ifchronos@timeline@showyears
\chronos@timeline@showyearstrue
\newif\ifchronos@eventyearsonline
\chronos@eventyearsonlinefalse
\pgfkeys{/pgf/number format,
  int detect,
  set thousands separator={},
}
\tikzset{%
  /handlers/.chronos too/.code={%
    \edef\chronos@tempc{\pgfkeyscurrentpath}%
    \edef\chronos@tempd{#1}%
    \tikzset{%
      \pgfkeyscurrentpath @too/.code={%
        \tikzset{%
          /chronos/\chronos@tempd/.append style={##1},
        }%
      },
      \chronos@tempc/.forward to=\chronos@tempc @too,
    }%
  },
  chronos/.code={% https://tex.stackexchange.com/a/159856/ - Claudio Fiandrino
    \tikzset{%
      align=center,
      anchor=mid,
      fixed point arithmetic,
      /chronos/.cd,
      #1,
    }%
    \pgfmathsetmacro\chronos@unit{(\chronos@width-20pt)/(\thechronos@enddate-\thechronos@startdate)}%
    \draw [/chronos/timeline@line, line width=\chronos@height] (-10pt,0) coordinate (chronos pre) -- +(\chronos@width,0) coordinate (chronos post);
    \coordinate (chronos base) at (0,-.5*\chronos@height);
    \coordinate (chronos top) at (0,.5*\chronos@height);
    \coordinate (chronos foot) at (0,{-.5*\chronos@height-\chronos@borderheight});
    \coordinate (chronos head) at (0,{.5*\chronos@height+\chronos@borderheight});
    \coordinate (chronos start) at (0,0);
    \coordinate (chronos end) at ([xshift=-10pt]chronos post);
    \ifdim\chronos@borderheight>0pt
      \fill [bottom color=chronos@borderinner, top color=chronos@borderouter] (chronos pre |- chronos head) rectangle (chronos post |- chronos top);
      \fill [top color=chronos@borderinner, bottom color=chronos@borderouter] (chronos pre |- chronos base) rectangle (chronos post |- chronos foot);
    \fi
    \ifchronos@timeline@showyears
      \pgfmathsetcounter{chronos@startyear}{\chronos@startyear}%
      \pgfmathsetcounter{chronos@endyear}{\chronos@endyear}%
      \def\tempa{none}%
      \ifx\chronos@stepfrom\tempa
        \def\tempa{01}%
        \ifx\chronos@startmonth\tempa
          \ifx\chronos@startday\tempa
            \else\stepcounter{chronos@startyear}%
          \fi
          \else\stepcounter{chronos@startyear}%
        \fi
      \else
        \pgfmathsetcounter{chronos@startyear}{\chronos@stepfrom}%
      \fi
      \def\tempa{12}%
      \def\tempb{31}%
      \ifx\chronos@endmonth\tempa
        \ifx\chronos@endday\tempb
          \stepcounter{chronos@endyear}%
        \fi
      \fi
      \pgfmathsetmacro\chronos@nextstep{int(\thechronos@startyear+\chronos@stepyears)}%
      \foreach \i in {\thechronos@startyear,\chronos@nextstep,...,\thechronos@endyear} {%
        \pgfcalendardatetojulian{{\i}-01-01}{\c@chronos@yeardate}%
        \pgfmathsetmacro\chronos@tempa{(\thechronos@yeardate-\thechronos@startdate)*\chronos@unit}%
        \def\tempa{on line}%
        \ifx\tempa\chronos@timelineyears
          \node (chronos@year@\i) [/chronos/.cd, timeline@years, timeline year on line] at (\chronos@tempa pt,0) {\i};
          \ifchronos@marks
             \path [/chronos/timeline mark on line] (chronos@year@\i.south) -- (chronos@year@\i |- chronos base);
             \path [/chronos/timeline mark on line] (chronos@year@\i.north) --  (chronos@year@\i |- chronos top);
          \fi
        \else
          \node (chronos@year@\i) [/chronos/.cd, timeline@years, timeline year off line] at (\chronos@tempa pt,0) {\i};
          \ifchronos@marks
            \path [shorten <=.5*\chronos@height, /chronos/timeline mark off line] (\chronos@tempa pt, 0) -- (chronos@year@\i);
          \fi
        \fi
      }%
    \fi
    \ifchronos@eventyearsonline
    \tikzset{%
      /chronos/timeline years=on line,
    }%
    \fi
  },
  chronos set date/.code args={#1:#2:#3:#4}{%
    \pgfcalendardatetojulian{{#1}-#2-#3}{\c@chronos@date}%
    \setcounter{chronos@#4date}{\thechronos@date}%
    \expandafter\def\csname chronos@#4year\endcsname{#1}%
    \expandafter\def\csname chronos@#4month\endcsname{#2}%
    \expandafter\def\csname chronos@#4day\endcsname{#3}%
  },
  chronos date/.style args={#1-#2-#3}{%
    /tikz/chronos set date/.expanded={#1:#2:#3:thing}%
  },
  chronos period date/.style args={#1-#2-#3}{%
    /tikz/chronos set date/.expanded={#1:#2:#3:otherthing}%
  },
  /chronos/.search also={/tikz},
  /chronos/.cd,
  step years/.store in=\chronos@stepyears,
  step from year/.store in=\chronos@stepfrom,
  start date/.style args={#1-#2-#3}{%
    /tikz/chronos set date/.expanded={#1:#2:#3:start},
  },
  end date/.style args={#1-#2-#3}{%
    /tikz/chronos set date/.expanded={#1:#2:#3:end},
  },
  timeline width/.store in=\chronos@width,
  timeline height/.store in=\chronos@height,
  width/.forward to=/chronos/timeline width,
  height/.forward to=/chronos/timeline height,
  timeline border height/.store in=\chronos@borderheight,
  timeline border inner colour/.code={\colorlet{chronos@borderinner}{#1}},
  timeline border outer colour/.code={\colorlet{chronos@borderouter}{#1}},
  timeline years set/.store in=\chronos@timelineyears,
  timeline years/.is choice,
  timeline years/.forward to=/chronos/timeline years set,
  timeline years/above/.code={%
    \tikzset{%
      /chronos/timeline@years/.style={above, anchor=south, yshift=.5*\chronos@height},
    }%
  },
  timeline years/below/.code={%
    \tikzset{%
      /chronos/timeline@years/.style={below, anchor=north, yshift=-.5*\chronos@height},
    }%
  },
  timeline years/on line/.code={%
    \tikzset{%
      /chronos/timeline@years/.style={anchor=center},
    }%
  },
  only years/.code={%
    \chronossetdateformat{Y}%
    \tikzset{%
      /tikz/chronos date/.style={%
        /tikz/chronos set date/.expanded={##1:01:01:thing}%
      },
      /tikz/chronos period date/.style={%
        /tikz/chronos set date/.expanded={##1:01:01:otherthing}%
      },
    }%
  },
  only text/.code={%
    \tikzset{/chronos/only years}%
    \chronossetdateformat{}%
  },
  background/.code={%
    \colorlet{chronos@background}{#1}%
  },
  timeline marks/.is if=chronos@marks,
  timeline show years/.is if=chronos@timeline@showyears,
  timeline no years/.code={%
    \tikzset{%
      /chronos/timeline show years=false,
    }%
  },
  lines/.style={draw, {Triangle[width=0pt 3,reversed,length=0pt 1.5]}-{Triangle[width=0pt 3,reversed,length=0pt 1.5]}, shorten <={.5*\chronos@height}},
  events/.style={fill=chronos@background, fill opacity=.95, text opacity=1, draw opacity=1, rounded corners, align=center, font=\footnotesize},
  period/.style={draw},
  period event line/.style={/chronos/lines},
  period event/.style={/chronos/events},
  event line/.style={/chronos/lines},
  event/.style={/chronos/events},
  event years on line/.is if=chronos@eventyearsonline,
  event year on line/.style={/chronos/timeline@years, /chronos/timeline year on line},
  special date/.store in=\chronos@specialdate,
  timeline@line/.style={draw},
  timeline line/.chronos too=timeline@line,
  timeline year off line/.style={fill=chronos@background, text opacity=1, fill opacity=.95},
  timeline mark off line/.style={draw, {Triangle[width=0pt 3,reversed,length=0pt 1.5]}-, thin, shorten >=-2pt},
  timeline year on line/.style={text=chronos@background, inner sep=1pt},
  timeline mark on line/.style={draw=chronos@background, shorten >=1.5pt},
  timeline mark too/.code={%
    \tikzset{%
      /chronos/.cd,
      timeline mark on line/.append style={#1},
      timeline mark off line/.append style={#1},
    }%
  },
  timeline year too/.code={%
    \tikzset{%
      /chronos/.cd,
      timeline year on line/.append style={#1},
      timeline year off line/.append style={#1},
    }%
  },
  timeline mark/.forward to=/chronos/timeline mark too,
  timeline year/.forward to=/chronos/timeline year too,
  start date=1001-10-01,
  end date=1003-06-14,
  timeline width=100mm,
  timeline height=1pt,
  timeline border height=0pt,
  chronos date=1850-01-01,
  chronos period date=1851-01-01,
  step years=1,
  timeline years=above,
  background=white,
  timeline border inner colour=black,
  timeline border outer colour=chronos@background,
  step from year=none,
  special date=none,
}
\NewDocumentCommand \chronosevent { O {} m O {} +m D () { -10pt } }
{%
  \tikzset{%
    chronos date/.expanded={#2},
  }5
  \pgfmathsetmacro\chronos@offset{(#5 < 0pt) ? (#5-.5*\chronos@height-.5*\chronos@borderheight) : (#5+.5*\chronos@height+.5*\chronos@borderheight)}%
  \pgfmathsetmacro\chronos@anchor{(#5 < 0pt) ? "north" : "south"}%
  \scoped[on background layer]{\path [postaction={/chronos/event line, #1}] ({(\thechronos@thingdate-\thechronos@startdate)*\chronos@unit pt},0) -- +(0,\chronos@offset pt) node [anchor=\chronos@anchor, /chronos/event, #3] {\chronos@showdate{thing}\\#4};}
  \ifchronos@eventyearsonline
    \edef\tempa{none}%
    \edef\tempb{\chronos@specialdate}%
    \ifx\tempa\tempb\else\let\chronos@thingyear\chronos@specialdate\tikzset{/chronos/special date=none}\fi
    \node [/chronos/.cd, event year on line] at ({(\thechronos@thingdate-\thechronos@startdate)*\chronos@unit pt},0) {\chronos@thingyear};
  \fi
}
\NewDocumentCommand \chronosspecialdate { m }
{%
  \tikzset{%
    /chronos/special date=#1,
  }%
}
\NewDocumentCommand \chronosperiod { O {} m O {} m O {} +m D () { -10pt } }
{%
  \tikzset{%
    chronos date/.expanded={#2}, chronos period date/.expanded={#4}
  }%
  \pgfmathsetmacro\chronos@offset{(#7 < 0pt) ? (#7-.5*\chronos@height-.5*\chronos@borderheight) : (#7+.5*\chronos@height+.5*\chronos@borderheight)}%
  \pgfmathsetmacro\chronos@anchor{(#7 < 0pt) ? "north" : "south"}%
  \path [postaction={line width=\chronos@height, /chronos/period, #1}] ({(\thechronos@thingdate-\thechronos@startdate)*\chronos@unit pt},0) -- ({(\thechronos@otherthingdate-\thechronos@startdate)*\chronos@unit pt},0);
  \scoped[on background layer]{\path [postaction={/chronos/period event line, #3}] ({(.5*\thechronos@otherthingdate+.5*\thechronos@thingdate-\thechronos@startdate)*\chronos@unit pt},0) -- +(0,\chronos@offset pt) node [anchor=\chronos@anchor, /chronos/period event, #5] {\chronos@showdate{thing}--\chronos@showdate{otherthing}\\#6};}
}
\makeatother

\DeclareCiteCommand{\eventcite}{}{\chronosevent{\thefield{year}}{\printnames{labelname}}}{}{}

@@

\newcommand{\onlyinsubfile}[1]{#1}
\newcommand{\notinsubfile}[1]{}
\begin{document}
\renewcommand{\onlyinsubfile}[1]{}
\renewcommand{\notinsubfile}[1]{#1}

\begin{frame}
\titlepage
\end{frame}
% \setbeamercovered{transparent}

\subfile{introduction.tex}

\begin{frame}{A case for comparative historical analysis}
\begin{itemize}
  \item Historical analysis isn't new
  \begin{itemize}
    \item In general

    \scriptsize{
      \textcite{rosenberg1994exploring,rosenberg1982inside}
    }
    \item In HCI

    \scriptsize{
      \textcite{Wyche2006,bodker1993historical}
    }
  \end{itemize}
  % \item Still can be useful.
\end{itemize}
\end{frame}
